BIOL 274 Homework 1

### 2024-01-20 bio_dynamics/bio274_hw1.Rmd
finished hw1

We are given $\frac{x}{x_0}=0.5$ at $t=30$ hours. Plugging into our equation, we get
$$0.5=e^{30k}$$
$$\ln(0.5)=\ln(e^{30k})$$
$$\ln(0.5)=30k$$
$$k=\frac{\ln(0.5)}{30}$$
```{r solve k}
k = log(0.5)/30
k
```

@@

```{r solve t}
t = log(0.01)/k
round(t, 0)
```

@@

### a
$$ \frac{dy}{dt}=\frac{3t+1}{2y}$$
$$ 2y \, dy=(3t+1)dt$$
$$\int 2y \, dy=\int(3t+1)dt$$
$$y^2=\frac{3t^2}{2}+t+C$$
$$y=\sqrt{\frac{3t^2}{2}+t+C}$$

### b
$$\frac{dy}{dx}=x\sqrt[3]{y}$$

@@

### c
$$\frac{dp}{dt}=\frac{1}{3tp^2}$$
$$3p^2\,dp=\frac{1}{t}dt$$
$$\int 3p^2\,dp=\int \frac{1}{t}dt$$
$$p^3=\ln(t)+C$$
$$p=\sqrt[3]{\ln(t)+C}$$

>>>>

We are given $\frac{x}{x_0}=0.5$ at $t=30$ hours. Plugging into our equation, we get
$$0.5=e^{30k}$$
$$\ln(0.5)=\ln(e^{30k})$$
$$\ln(0.5)=30k$$
$$k=\frac{\ln(0.5)}{30}$$
```{r solve_k_1}
k = log(0.5)/30
k
```

@@

```{r solve_t_1}
t = log(0.01)/k
round(t, 0)
```

@@

### a)
$$ \frac{dy}{dt}=\frac{3t+1}{2y}$$
$$ 2y \, dy=(3t+1)dt$$
$$\int 2y \, dy=\int(3t+1)dt$$
$$y^2=\frac{3t^2}{2}+t+C$$
$$y=\sqrt{\frac{3t^2}{2}+t+C}$$

### b)
$$\frac{dy}{dx}=x\sqrt[3]{y}$$

@@

### c)
$$\frac{dp}{dt}=\frac{1}{3tp^2}$$
$$3p^2\,dp=\frac{1}{t}dt$$
$$\int 3p^2\,dp=\int \frac{1}{t}dt$$
$$p^3=\ln(t)+C$$
$$p=\sqrt[3]{\ln(t)+C}$$

## Question 4
>According to Newton’s law of cooling, the rate of decrease of temperature of a body is proportional to the difference between its temperature and that of its environment. If the temperature in your living room is 20°C, you remove a log from the fireplace at 100°C, and it takes 10 minutes to cool to 60°C, how long will it take to decrease the temperature of the log to 25°C?

$$\frac{dT}{dt}=k(T_{body}-T_{env})=k(T_{body}-20)$$
$$
\frac{1}{(T_{body}-20)}dT=k\,dt
$$
$$\int\frac{1}{(T_{body}-20)}dT=kt+C$$

$$\ln{(T_{body}-20)}=kt+C$$
$$T_{body}=e^{kt+C}+20=Ce^{kt}+20$$

Since $T_{body}=C+20$ at time $t=0$, we can write the equation as
$$T_{body}(t)=(T_0-20)e^{kt}+20$$
for the temperature of the body at time $t$.
Now using the information we are given,
$$60=(100-20)e^{10k}+20$$
$$40=80e^{10k}$$
$$\frac{1}{2}=e^{10k}$$
$$\frac{\ln{(1/2)}}{10}=k$$

```{r solve_k_2}
k2 = log(1/2)/(10)
k2
```

Now we want to solve for $t$ in the following equation:
$$25=(60-20)e^{kt}+20$$
$$5=40e^{kt}$$
$$\frac{\ln{(1/8)}}{k}=t$$

```{r solve_t_2}
t2 = (log(1/8))/(k2)
t2
```

*Quick sanity check*:
If it took 10 minutes to cool from 100°C to 60°C and 30 minutes to cool from 60°C to 25°C, it should take 40 minutes to cool from 100°C to 25°C:
```{r}
# 25=(100-20)e^kt+20
# 5=80e^kt
# ln(1/16)=kt

t3=(log(1/16))/(k2)
t3
```

## Question 5
Solve each of the following IVPs:

### a)
$$\frac{dx}{dt}=-xt,\, x(0)=1/\sqrt{\pi}$$

$$\frac{1}{x}dx=-t\,dt$$
$$\int \frac{1}{x}dx=\int-t\,dt$$
$$\ln{(x)}=-\frac{t^2}{2}+C$$

$$x(t)=Ce^{-t^2/2}$$
is our general solution to the differential equation.
Now we plug in the initial conditions:
$$1/\sqrt{\pi}=Ce^{-(0^2/2)}=Ce^0=C$$

$$x(t)=\frac{e^{-t^2/2}}{\pi}$$

is the solution to the IVP.

### b)
$$\frac{dy}{dx}=xy^2+2y^2,\, y(0)=1$$

$$\frac{dy}{dx}=y^2(x+2)$$
$$y^{-2}dy=(x+2)dx$$

$$-\frac{1}{y}=\frac{x^2}{2}+2x+C$$

We can plug in our initial conditions now:
$$-\frac{1}{1}=0+0+C$$
so $C=-1$.
$$-\frac{1}{y}=\frac{x^2}{2}+2x-1$$
$$\frac{1}{y}=1-\frac{x^2}{2}-2x$$
$$\frac{1}{y}=1-\frac{x^2}{2}-2x$$
or
$$\frac{2}{y}=2-x^2-4x$$

$$y(x)=\frac{2}{2-x^2-4x}$$
is the solution to the IVP.

### c)

$$ \frac{du}{dt}=\frac{\cos(t)}{9u^2}, \, u(0)=2$$

$$ u^2du=\frac{\cos(t)}{9}dt$$
$$ \int u^2du=\frac{1}{9}\int \cos(t)dt$$

$$ \frac{u^3}{3}=\frac{1}{9}\sin(t)+C$$

$$ u^3=\frac{\sin(t)}{3}+C$$

Now we plug in the initial conditions:
$$ 2^3=\frac{\sin(0)}{3}+C=0+C$$
So $C=8$ and we have
$$ u(t)=\sqrt[3]{\frac{\sin(t)}{3}+8}$$

as the solution to the IVP.
